```{r}
#definition
```

```{r}
reverse <- function(input) {
 return( unlist(input)[length(input):1] )
}
```

```{r}
reverser <- function(x){
Alpha <- strsplit(x, split= NULL)
Beta <- rev(Alpha[[1]])
return(paste(Beta,collapse=""))
}
```

```{r}
scrambler <- function(x){
Alpha <- strsplit(x, NULL)
Beta <- sample(Alpha[[1]])
return(paste(Beta,collapse=""))
 }
```

```{r}
VowelBleeper <- function(input) {
supersub <- gsub("[aeiou]","*",input)
return(supersub)
}
VowelBleeper("dog")
```

```{r}
Cypher <- function(input) {
sub <- gsub("[e]","3",input)
sub2 <- gsub("[o]","0",sub)
sub3 <- gsub("[s]","5",sub2)
sub4 <- gsub("[g]","9",sub3)
return(sub4)
}
Cypher("go")
```

```{r}
L33t<- function(input) {
sub <- chartr("e","3",input)
sub2 <- chartr("o","0",sub)
sub3 <- chartr("s","5",sub2)
sub4 <- chartr("g","9",sub3)
return(sub4)
}
L33t("go")
```

reserved words= NA Inf NaN NULL TRUE FALSE function
grep gsub regexp nchar length strsplit unlist factorial

```{r}
myfactorial <- function(n) { 
  res <- 0
  for(k in 0:n){
  res <- res + k
  }
  return(res)
  }
```

findScrable <- function(letters) {
for (k in 1:length(letters)) {
words <- words [grep(letters[k],words)]
if(length(words)<10) break
}
return(words)
}
findScrable(c("b","n","r","x","a","c","z","q"))
 
fib <- function(n) { 
current <- 1
beforeIt <- 0
for(k in 1:n) {
tmp <- current + beforeIt 
beforeIt <- current 
current <- tmp
}
return(current)
}

fib <- function(n) {
if( n < 0 | round(n) !=n){
return(NA)
}
sofar=c(0,1)
for(k in 3:n) {
sofar[k] <- sofar[k-1] + sofar [k-2]
}
return(sofar)
}

guess <- rep(TRUE, 100)
for(k in 2:100)
guess[(1:100)%%k == 0&(1:100)>k] <- FALSE
}
return(which(guess))

primeSieve<- function(n) {
guess <- rep(TRUE,n)
inds<- 1:n
for (k in 2:n) {
guess[(ind%%k)== 0 & inds>k)] <- FALSE
}
return(which(guess))
}

pigLatinRule <- function(word){
res <- 0
if (grep1"[^aeiou.+$]", word)) {
res <- "No Vowels"
}
break
if(grep1(^aeiou+$, word){
res <- "Consonant Start"
else{
res <- "Vowel Start"}
return(res)
}

Rule 1 <- funvtion(word)

syntax structure 
for while repeat 

for (k in 1:something)
function(,,){something}

while(logical x){changes x}
repeat{something, there needs to be a break}

looping algorithm 
-> setting up a state 
-> loop- state update
       - termination 
-> do something with the state

mySum <- function(v) {
         sofar<-0 
         k<-1
         for (sofar <- sofar+ v[k]){
         k <- k+1
         if (k==length(v)) break
         }
         return(sofar)
}

mySum <- function(v){
         sofar <- 0
         k <- 1
         repeat{
         sofar <- sofar + v[k]
         k <- k + 1
         
         if
         (k== (length(v) + 1)) break
         }
         return(sofar)
}

mySumWhile <- function(v){
              sofar <- 0
              k <- 1
              while
              (k !<= (length(v)+1) )
              
              sofar <- sofar + v[k]
              k <- k + 1
              
              return(sofar)
           }

mySumFor <- function(v) {
            sofar <- 0
            for(k in 1: length(v))
            sofar <- sofar + v[k]
            
            return(sofar)
}

myRunningSum <- function(v) {
                sofar <- 0
                res <- c()
                for (k in 1: length(v)) {
                sofar <- sofar + v[k]
                res <- c(res, sofar)
                }
                return(res)
               }
               
myRunningSum <- function(v) {
                sofar <- 0
                res <- c()
                for (k in v) {
                sofar <- sofar + k
                res <- c(sofar, res)
                }
                return(res)
               }
               
 
myUnique <- function(v) {
            already <- c()
            for (k in v) {
            if (k %in% already) {
            already <- c(already,"***")
            } else{
            already <- c(already,k)
          }
          }
          return(already)
          }
```{r}
toBase <- function(Z,b) {
          sofar <- c()
          repeat{
          r <- Z%%b
          sofar <- c(sofar,r)
          Z <- (Z-r)/b

          if (Z==0) break
          }
          
          return(sofar)
          }
```          
Given a character vector of digits and a numerical base

```{r}
baseToNumeric <- function(Nvec,b) {
  howMany <- length(Nvec)
  herdsize <- b^((howMany-1):0)
  return(sum(as.numeric(Nvec)*herdsize))
} 

```

### As a loop
```{r}
convertAsALoop <- function(Nvec,b) {
#Initial state
sheepCount <- 0
boxSize <- 1

for( k in length(Nvec):1){
 sheepCount <- sheepCount + boxSize*Nvec[k]
 boxSize <- boxSize*b 
 }
 
 return(sheepCount)
}
```
 
 
```{r}
blastoffWhile <- function(time){
       while(time!=0){
         Sys.sleep(1)
         cat(time,"")
         time <- time -1 
       }
       Sys.sleep(1)
       cat("Blastoff!")
}

```


```{r}
blastoffWhile <- function(time) { # Time is a pos. integer in seconds
  while (time != 0){
    Sys.sleep(1)
    cat(time,"\n")
    time <- time -1
  }
  Sys.sleep(1)
  cat("Blastoff!!!!")
  }

```

```{r}
countdown <- function(k) {
   system("say five")
   Sys.sleep(1)
   system("say four")
   Sys.sleep(1)
   system("say three")
   Sys.sleep(1)
   system("say blastoff!")
}
```

```{r}
countdown <- function(time) {
 
   for (k in time:1) {
   starttime <- Sys.time()
   system(paste("say",k))
   now <- Sys.time()
   Sys.sleep(1 - as.numeric(now-starttime))
   }
   system("say blastoff!")
}
```

```{r}
getData <- function(N){
for(k in 1:10){
before <- Sys.time()
readline("Press return")
after <- Sys.time()
delay <- as.numeric(after-before)
cat(rep("\n",20))
result[k] <- delay
Sys.sleep(runif(1,min=1.max=5))
     }
return(result)
 }
```

-object names
 x<- 7
 sqrt(x)
-operators/function
name(3,7,col="red")
==()
-arguments
-control structures
loops: for while repeat if(){ } else { } break continue
-function
packages up, myFun<- function(x,y,z=1){...}
-kinds of things
 class(thing) vectors lists dataframe numeric charaterstrings time
-indexing 
 v[ ] v[1:3] v[[]] [-1] v[c(TRUE,FALSE,FALSE,..)]

environment
list(a=1, b="hello")

y <-"hello"
with(globalenv(),y)

ls(globalenv())

y <- "hello"
get("y",globalenv())
assign("y",90,globalenv())
assign("44 skadoo",10,globalenv())

ls(baseenv())

sin <- function(x){
       3*x + 2
}
sin(pi)

remove("sin")

current environment <- function invocation environment 
global environment <-base environment <- 


f(x=2)
f <- function(x){
g(x+2)
}

g <- function(x){
 x^3
}


foo <- 4
f <- function(x) {
 foo <- 3
 g(7)
}
g <- function(x) { 
  foo^3
}

addEm <- function(v){
sum <- 0
for (k in 1:length(v)){
sum <- sum + v[k]
}
return(sum)
}
addEm(1:3)

# Reverse a vector, recursively

```{r}
revrec <- function(v) {
  if (length(v)==1) return(v)
  else {
    c(revrec(v[-1]), v[1])
  }
  }
revrec(c(5,4,3,2,1))
```

```{r}
revrec <- function(v) {
  browser()
  if (length(v)==1) return(v)
  else {
    c(revrec(v[-1]), v[1])
  }
  }
revrec(c(5,4,3,2,1))
```

searching can be done in 'log base 2(N)'


function(v){
if (length(v)<= 1) return(v)
mid <- v[ceiling(runif(1,min=2),max=length(v)))]
mids <- v[v==mid]
left <- v[lessthan(v,mid)]
right <- v[lessthan(mid,v)]

}

```{r}
myD <- function(f){
  h <- 0.001
  fprime<- function(x) {(f(x+h)- f(x-h))/h} 
  return(fprime)
   
}
```

```{r}
rossD <- function(f,x) {
      h <- 0.001
      res <- (f(x+h)-f(x))/h
      return(res)
}
```

```{r}
myD <- function(f, h=0.0000001 ) {
  fprime <- function(x) {(f(x+h)-f(x))/h}
  return(fprime)
}


myD2 <- function(f,h=0.0000001){ 
  gprime <- myD(f, h=h)
  eprime <- myD(gprime, h=h)
  return(eprime)
   
}
```

landscape
651-295-7487

```{r}
readImageURL <- function(URL){
  isJPEG <- grepl("[.jpg$|.jpeg$]",URL)
  
  if( isJPEG ) 
   image <- readJPEG(getURLContent(URL))
    else
   image<- readPNG(getURLContent(URL))
      
  }
```

##Plan for the next couple of weeks
GUI
manipulate -> shiny
- one plot
-Rstudio
-Control layout limited
-Run within R

shiny
-embed in document
-arbitrary layouts
-multiple outplots
 plot text table image
-Any R
-Without any R
Data and graphics
database operation machine graphics-> maps
R -> html
```{r}
x <- seq(0,1, length=101)
plot(1:2, ylim= c(0,5), xlim= c(0,1), xlab= "x", ylab= "Prob Density",)
 plot(1:2, ylim= c(0,5), xlim= c(0,1), xlab= "x", ylab= "Prob Density",type= "n")
 lines(x, dbeta(x,shape1=1,shape2=2), lwd=1)
 lines(x, dbeta(x,shape1=1, shape2=4),lwd=2)
 lines(x, dbeta(x,shape1=4, shape2=2),lwd=3)
```

```{r}
drawBeta <- function(a,b,...){ 
plot(1:2, ylim= c(0,5), xlim= c(0,1), xlab= "x", ylab= "Prob Density",)
 plot(1:2, ylim= c(0,5), xlim= c(0,1), xlab= "x", ylab= "Prob Density",type= "n")
lines(x, dbeta(x,a,b),...)
}  
```

```{r eval=FALSE}
betaApp <- function(pts) {
manipulate( {drawBeta(a,b,col=color);
             points(pts,rep(0,length(pts)))},
           a=slider(0,10,initial=1, label="a", step=.2),
             b= slider(0,10,initial=2, label="BEE", step=.2),
           color= picker(list("red","blue","green"))
            )
}
```

```{r}
animals <- list("Animals", list("Domesticated",list("Horse","Cat","Dog")), list("Wild", list("Lion","Tiger","Bear","Moose")))

displayTree(animals)
```

```{r}
countries <- ourTree <- list("The World",
  			list("Non-Continental", list("Fiji","New Zealand")),
				list("Continental",
               	  	  list("Africa", list("Ghana", "Kenya", "Morocco")),
                      list("Americas",
                      		list("North", list("Canada","Mexico")),
                            list("South",list("Chile","Bolivia"))),
                      list("Antarctica"),
                      list("Asia", 
                      		list("Thailand", "Bhutan", "Mongolia")),
                      list("Australia"),
					  list("Europe",
                    		list("Slovakia", "Romania"))))
displayTree(countries)
```

```{r}
myUI <- bootstrapPage( 
  div(class='jumbotron masthead', 
      h1('The Title of My Page'),
      p('This will be the very polite introduction to my page')
      ), 
  div(class='row span12',
      
    div(class='span4',
        h3('In the first block')
        ),
    div(class='span3',
        h4('I am here, too'),
        h2('Big Brother')
        )
    )
  )
  
```

```{r}
x<- seq(0,1, length =101)
plot(1, ylim = c(0,5), xlim = c(0, 1), xlab = "x", ylab = "Prob Density",
   type="n")
lines(x, dbeta(x, 1, 2), lwd =1)
lines(x, dbeta(x, 1, 4), lwd =2)
lines(x, dbeta(x, 4, 2), lwd =3)

betaDist<- function(a, b, ...){
  x<- seq(0,1, length =101)
  plot(1, ylim = c(0,5), xlim = c(0, 1), xlab = "x", ylab = "Prob Density",
     type="n")
  lines(x, dbeta(x, a, b), ...)
}

require(manipulate)

betaApp<- function(pts){
manipulate( {betaDist(a,b, col=r, lwd=t);
             points(pts, rep(0,length(pts)))},
            a =slider(0,10, initial=1, label="a", step=.2),
            b =slider(0,10, initial=2, label= "b", step=.2),
            r = picker("black",Ruby="red",OrcSkin="green",SpringWater="blue",
                       Yellow="yellow", initial = NULL, label="Color"),
            t =slider(0,4, initial=1.5, label= "Line Thickness", step=.5)
    )
}
```


diacolors <- with( small, ifelse(diab==0, alpha("blue",.2), alpha("red",.4)))

diasize <- with(small, ifelse(diab==0, 1,3))

plot(wgt~hgt, data=small, col=diacolors, cex=diasize, pch=20)

computing
computation from inputs to outputs
change in state from response to inputs

sum(vector/numeric/logical)
%% == + - * ^
function invocation
f(   )

function(arg list){expression}

function
is.majority(votes)
what.percentage(votes)
name # of votes  percent votes




